Bare minimum: set theory

### 2021-05-19 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: adding references

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

@@

TODO: cite https://www.math24.net/topics-set-theory and set theory book

@@

A \textbf{class} \marginright{TODO: diagram with Classes, proper classes and sets.}  is any collection, therefore \textbf{sets are classes}. Some classes are not sets. These are called \textbf{proper classes} and are exactly the collections that cannot be elements of a class. Therefore, sets are classes that can be contained by other classes, while proper classes cannot. The difference between sets and proper classes is not in what they contain but in whether they can be contained. Proper classes are ``too large'' to be sets in the following sense: given a class $A$, if we can find a set $B$ that contains all elements of $A$, then $A$ will be a set as well. Therefore proper classes can never be subclasses of sets, they are bigger than all sets. The class of all sets, for example, is a proper class. So is the class of all cardinal numbers.

@@

\textbf{Axioms of set theory}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE that follows NBG:
\begin{enumerate}
	\item Axiom of extent. If two classes have the same elements, then they are identical.
	\item Axiom of class construction. Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements (i.e. sets) x which satisfy $P(x)$.
	\item Every subclass of a set is a set.
	\item $\emptyset$ is a set.
	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
\end{enumerate}

@@

%\bibliographystyle{plain}
%\bibliography{bibliography}

>>>>

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

@@

For more details on set theory, see for example \cite{settheory-tutorial, pinter2014book}.

@@

A \textbf{class} \marginright{ ~\\
		\begin{tikzpicture}[scale = 0.3]
		\node [draw, inner sep=6, rectangle] (class) {\small Class};
		\node [below=.3 of class] (middle) {};
		\node [below=1.3 of class] (lower) {};
		\node [draw, inner sep=6, rectangle, align=center, left=0 of lower] (pclass) {\small Proper \\ class};
		\node [draw, inner sep=6, rectangle, right=.2 of lower] (set) {\small Set};
		\draw[->] (pclass.north) -| ++(0,.8) -|  (class.south);	
		\draw[->] (set.north) -| ++(0,1.5) -|  (class.south);	
		\draw[->] (class.east) -| ++(3,0) |-  (set.east);	
		 %++(0,-1) -|
		\node [left=-0.1 of middle] (isa) {\small is a};
		\node at (4.1,.6) (contains) {\small contains};
	\end{tikzpicture} 
}  is any collection, therefore \textbf{sets are classes}. Some classes are not sets. These are called \textbf{proper classes} and are exactly the collections that cannot be elements of a class. Therefore, sets are classes that can be contained by other classes, while proper classes cannot. The difference between sets and proper classes is not in what they contain but in whether they can be contained. Proper classes are ``too large'' to be sets in the following sense: given a class $A$, if we can find a set $B$ that contains all elements of $A$, then $A$ will be a set as well. Therefore proper classes can never be subclasses of sets, they are bigger than all sets. The class of all sets, for example, is a proper class. So is the class of all cardinal numbers.

@@

\textbf{Axioms of set theory}. Here \marginleft{Set theory axioms} are the axioms of a particular formulation of axiomatic set theory taken from \cite{pinter2014book} that follows NBG:
\begin{enumerate}
	\item Axiom of extent. If two classes have the same elements, then they are identical.
	\item Axiom of class construction. Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements (i.e. sets) x which satisfy $P(x)$.
	\item Every subclass of a set is a set.
	\item $\emptyset$ is a set.
	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
\end{enumerate}

@@

\bibliographystyle{plain}
\bibliography{bibliography}
